Analysis on Trend of Death of Shelter Residence in Toronto 

### 2024-09-25 paper/paper.qmd
add more cross-reference chang part of the paper text content

**Data Quality and Cleaning**: The dataset required careful cleaning to address issues such as missing values and inconsistencies, particularly in the reporting of deaths by gender. Although high-level cleaning was conducted to ensure the reliability of the data, the absence of certain variables may still introduce biases or limit the interpretability of the results.

@@

### Trend Over Time

@fig-shelter-death-line, created by @citeR, shows the total number of deaths of Toronto shelter residents each year from January 2007 to August 2024, divided into 15 sub-graphs, arranged in 5 rows and 3 columns by year. This layout allows readers to clearly compare the trend of deaths from one year to the next and observe potential seasonal changes.

@@

The purpose of this visualization is to analyze whether certain months have higher mortality rates. It opens the door to exploring potential seasonal influences on mortality rates, such as the impact of extreme temperatures or adverse weather conditions, which could contribute to increased mortality among shelter residents.

The current findings, shown in the line graph, indicate that January and November have the highest mortality rates, while March and September have the lowest mortality rates. These findings highlight the need for further research into the factors influencing these monthly variations, which could inform targeted interventions to improve the health and safety of shelter residents.

>>>>

**Data Quality and Cleaning**: The dataset misses values of transgender/Non-binary/two-spirit decedents until 2020 due to lack of awareness of gender identification. The missing values only affect the research in gender vulnerability field and it is hard to be fulfilled. In our research, the missing value is replaced with 0 mostly. The value keeps being NA when calculate the death ratio of each gender.

@@

### Trend Over Time {#sec-time}
@fig-shelter-death-line, created by @citeR, shows the total number of deaths of Toronto shelter residents each year from January 2007 to August 2024, divided into 15 sub-graphs, arranged in 5 rows and 3 columns by year. This layout allows readers to clearly compare the trend of deaths from one year to the next and observe potential seasonal changes.

@@

The purpose of this visualization is to analyze whether certain months have higher mortality rates. It helps to exploring potential seasonal influences on mortality rates, such as the impact of extreme temperatures or adverse weather conditions, which could contribute to increased mortality among shelter residents.

The current findings, shown in the line graph, indicate that January and November have the highest mortality rates, while March and September have the lowest mortality rates, which was different in what we find in @sec-time. These findings highlight the need for further research into the factors influencing these monthly variations, which could inform targeted interventions to improve the health and safety of shelter residents.
